---
title: "Human activity analysis"
author: "Sahil"
date: "24-oct"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview
In this project, my goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: (http://groupware.les.inf.puc-rio.br/har).

# Getting Data

```{r getting_data}
dtrain <- read.csv("C:/Users/sahil/Documents/pml-training.csv")
dtest <- read.csv("C:/Users/sahil/Documents/pml-testing.csv")
```

# Cleaning Data
As we can observe the data has a lot of columns having NA values and blank data. Let us remove these columns first.
```{r cleaning_data}
# Removing columns with blank values and NA values
r <- which(sapply(dtrain, function(x){sum(x=="")==0}))
dtrain <- dtrain[,r]
r <- which(sapply(dtest, function(x){sum(x=="")==0}))
dtest <- dtest[,r]
```

Now we should also remove the unimportant columns. The first 7 columns of the data frame would not help in a qualitative analysis.

```{r cleaning_data2}
imp_col <- c(8:dim(dtrain)[2])
dtrain <- dtrain[,imp_col]
dtest <- dtest[,imp_col]
```

Now let us convert the classes column as factor variables.

```{r cleaning_data3}
dtrain$classe <- as.factor(dtrain$classe)
```

# Cross Validation

```{r cross_validation}
library("caret")
intrain <- createDataPartition(dtrain$classe, p=0.60)
dcv <- dtrain[-intrain[[1]],]
dtraining <- dtrain[intrain[[1]],]
```

# Model Selection
Before doing any exploratory data analysis let us first apply certain models to see how they perform.
Let us apply a random forest model.

```{r model_selection}
library("randomForest")
rf_fit <- randomForest(classe~., data=dtraining, ntree=100,
                       do.Trace=T)
rf_fit
```

As we can see the error rate is less than 1%, we will stick with this model.

# Prediction
```{r}
pred <- predict(rf_fit, dtest)
print(pred)
```

# Conclusion

Hence, we have used a random forest model, after certain data cleaning to construct our predictor.